A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-19 manuscript.Rmd

While our approach was designed to scale well with the complexity of a computationally intense project, we realize that this flexibility may not be straightforward to integrate into researchers' everyday workflow. There are various R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow, and we recommend it to users who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analyses with low computational demands because binder's memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not guarantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited. A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs; @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with renv and dynamic document generation with R Markdown. We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects. It achieves a high standard of reproducibility but does not guarantee full computational reproducibility and is limited to dependency management within the R environment.

@@

How can one best share a reproducible workflow? We believe that, ideally, a non-commercial public service provider should be found that guarantees permanent and reliable hosting of our reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017]. A provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN would be desirable. Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the [Creative Commons license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT). Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work. These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009]. A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice. To facilitate an inclusive environment, we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.

@@

The proposed workflow leverages various existing tools that are partly integrated into RStudio already. Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]), which are particularly recommendable to beginners but they are either incomplete in the sense of our proposal or rely on proprietary service providers. We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools to allow reproducible workflows to become a standard in psychological research.

>>>>

While our approach was designed to scale well with the complexity of a computationally intense project, we realize that this flexibility may not be straightforward to integrate into researchers' everyday workflow. There are various R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow, and we recommend it to users who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analyses with low computational demands because binder's memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not guarantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited. A particularly noteworthy approach is the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs; @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with `renv` and dynamic document generation with R Markdown. We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects. It achieves a high standard of reproducibility but does not guarantee full computational reproducibility and is limited to dependency management within the R environment.

@@

How can one best share a reproducible workflow? We believe that, ideally, a non-commercial public service provider should be found that guarantees permanent and reliable hosting of our reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017]. A provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN would be desirable. Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format. Typically, there is no single license that works for code, data, and media (such as text or figures). We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place. In our experience, the [Creative Commons license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar). Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work. These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009]. A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice. To facilitate an inclusive environment, we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.

@@

The proposed workflow leverages various existing tools that are partly integrated into RStudio already. Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we particularly recommend to beginners;  but they are either incomplete in the sense of our proposal or rely on proprietary service providers. We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools and services to allow reproducible workflows to become a standard in psychological research.
